---
title: "Comparing DuckDB/Arrow Performance"
description: "How to evaluate memory and CPU usage for long-running processes in duckdb/arrow"
lang: en
date: 2025-07-10
categories: [R, benchmark]
image: https://duckdb.org/images/logo-dl/DuckDB_Logo-horizontal.svg
draft: false # setting this to `true` will prevent your post from appearing on your listing page until you're ready!
---

When it comes to comparing different approaches, the ideal scenario is to run the code in benchmarking tools, but the "classic" R tools are not well suited for comparing `duckdb` and/or `arrow` code:

- `tictoc` only returns elapsed time
- `bench` does not detect memory allocations from duckdb and arrow
- ...

In my articles, I will regularly use [timemoir](https://github.com/nbc/timemoir), written specifically for this type of comparison:

```{r}
#| label: timemoir
library(timemoir)

test_function <- function(n) {
  x <- rnorm(n); mean(x)
}

res <- timemoir(
  test_function(1.2e7),
  test_function(4e7),
  test_function(1e8)
)

res |> 
  kableExtra::kable()

plot(res)
```

---

That said, these are not "true" rigorous benchmarks—well beyond the scope of this blog, but rather quick comparisons intended to provide a rough idea of relative performance.

::: {.callout-note collapse=true}
## Session Information
```{r}
#| label: session_info
devtools::session_info(pkgs = "attached")
```
:::
